---
title: "Destilado de CSVs a tablas"
author: "arrozet"
editor: visual
---

# Carga de librerias, datasets y funciones de parseo
```{r setup_and_load_data}
# Cargar librerías necesarias
library(tidyverse)
library(jsonlite)

# Cargar los datasets (asegúrate que las rutas son correctas respecto a tu directorio de trabajo)
# Asumiendo que el Rmd se ejecuta con el directorio de trabajo en la raíz del proyecto MOMDB
raw_credits <- read_csv("./database/datasets/tmdb_5000_credits.csv")
raw_movies <- read_csv("./database/datasets/tmdb_5000_movies.csv")

# Función para parsear JSON de forma segura
parse_json_safely <- function(json_string) {
  tryCatch({
    if (is.na(json_string) || json_string == "" || json_string == "[]" || !jsonlite::validate(json_string)[1]) {
      # Devuelve un data frame vacío con estructura esperada si es posible, o NA genérico
      return(tibble()) # O NA, dependiendo de cómo quieras manejarlo después
    }
    fromJSON(json_string)
  }, error = function(e) {
    # message("Error parsing JSON: ", json_string, " - ", e$message) # Opcional para depuración
    return(tibble()) # O NA
  })
}
```

# Dataset Movies (películas y todo lo relacionado con ellas)
## Movie
```{r}
# 1. Procesar el dataframe de películas (raw_movies)

# Dataframe base de películas (sin las columnas JSON que se procesarán por separado)
movie <- raw_movies |>
  select(
    movie_id = id, # Renombrar id a movie_id para claridad
    original_title,
    budget,
    homepage,
    original_language,
    overview,
    popularity,
    release_date,
    revenue,
    runtime,
    status_name = status, # Se usará para crear la tabla Status
    tagline,
    title,
    vote_average,
    vote_count
  ) |>
  mutate(release_date = as.Date(release_date)) |>
  arrange(movie_id)

cat("--- Movies creado ---\n")
head(movie)
```

## Status
```{r}
# Tabla de Status únicos
status <- movie |>
  distinct(status_name) |>
  filter(!is.na(status_name)) |>
  mutate(status_id = row_number())

cat("\n--- Status creado ---\n")
head(status)
```

## Genres y Movie_Genre
```{r}
# Procesar Genres
movie_genre <- raw_movies |>
  select(movie_id = id, genres) |>
  mutate(parsed_data = map(genres, parse_json_safely)) |>
  filter(map_lgl(parsed_data, ~is.data.frame(.x) && nrow(.x) > 0)) |>
  unnest(parsed_data, names_repair = "unique") |>
  rename(
    genre_id = id,  # Rename the 'id' column (from JSON) to 'genre_id_json'
    genre_name = name    # Rename the 'name' column (from JSON) to 'genre_name'
  ) |>
  # Now columns are: movie_id, genres, genre_id_json, genre_name
  select(movie_id, genre_id, genre_name) |>
  arrange(movie_id, genre_id)

genre <- movie_genre |>
  distinct(genre_id, genre_name) |>
  arrange(genre_id)

movie_genre <- movie_genre |>
  select(-genre_name)

cat("\n--- Relacion Movies - Genre creada ---\n")
head(movie_genre)

cat("\n--- Genre creado ---\n")
head(genre)
```

## Keywords y Movie_Keywords
```{r}
# Procesar Keywords
movie_keywords <- raw_movies |>
  select(movie_id = id, keywords) |>
  mutate(parsed_data = map(keywords, parse_json_safely)) |>
  filter(map_lgl(parsed_data, ~is.data.frame(.x) && nrow(.x) > 0)) |>
  unnest(parsed_data, names_repair = "unique") |>
  rename(keyword_id_json = id...3, keyword_name = name) |>
  select(movie_id = movie_id...1, keyword_id_json, keyword_name)

keywords <- movie_keywords |>
  distinct(keyword_id_json, keyword_name)

cat("\n--- movie_keywords y keywords Creados ---\n")
head(movie_keywords)
head(keywords)
```

## ProductionCompany y Movie_ProductionCompany
```{r}
# Procesar Production Companies
movie_company <- raw_movies |>
  select(movie_id = id, production_companies) |>
  mutate(parsed_data = map(production_companies, parse_json_safely)) |>
  filter(map_lgl(parsed_data, ~is.data.frame(.x) && nrow(.x) > 0)) |>
  unnest(parsed_data, names_repair = "unique") |>
  rename(company_id = id, company_name = name) |>
  select(movie_id = movie_id, company_id, company_name) |>
  arrange(movie_id, company_id)

company <- movie_company |>
  distinct(company_id, company_name) |>
  arrange(company_id)

movie_company <- movie_company |>
  select(-company_name)

cat("\n--- movie_company y company Creados ---\n")
head(movie_company)
head(company)
```

## ProductionCountry y Movie_ProductionCountry
```{r}
# Procesar Production Countries
movie_country <- raw_movies |>
  select(movie_id = id, production_countries) |>
  mutate(parsed_data = map(production_countries, parse_json_safely)) |>
  filter(map_lgl(parsed_data, ~is.data.frame(.x) && nrow(.x) > 0)) |>
  unnest(parsed_data, names_repair = "unique") |>
  rename(country_iso = iso_3166_1, country_name = name) |>
  select(movie_id, country_iso, country_name) |>
  arrange(movie_id, country_iso)

country <- movie_country |>
  distinct(country_iso, country_name) |>
  arrange(country_iso)

movie_country <- movie_country |>
  select(-country_name)

cat("\n--- movie_country y country Creados ---\n")
head(movie_country)
head(country)
```

## SpokenLanguages y Movie_SpokenLanguages
```{r}
# Procesar Spoken Languages
movie_language <- raw_movies |>
  select(movie_id = id, spoken_languages) |>
  mutate(parsed_data = map(spoken_languages, parse_json_safely)) |>
  filter(map_lgl(parsed_data, ~is.data.frame(.x) && nrow(.x) > 0)) |>
  unnest(parsed_data, names_repair = "unique") |>
  rename(lang_iso = iso_639_1, lang_name = name) |>
  select(movie_id, lang_iso, lang_name) |>
  arrange(movie_id, lang_iso)

language <- movie_language |>
  distinct(lang_iso, lang_name) |>
  arrange(lang_iso)

movie_language <- movie_language |>
  select(-lang_name)

cat("\n--- movie_language y language Creados ---\n")
head(movie_language)
head(language)
```
Los lenguajes que salen con `lang_name` vacio es literalmente porque no tienen nombre en la base de datos de TMDB. Por ejemplo, `lang_iso = "bo"` está en la película con `original_title = "2012"`, y en `name` tiene la cadena vacía.

# Dataset Credits (toda la parte de Crew, Person...)

## Preprocesamiento
```{r}
# Process the 'crew' JSON column
credits_crew_raw <- raw_credits |>
  select(movie_id, crew_json = crew) |>
  mutate(parsed_crew = map(crew_json, parse_json_safely)) |>
  filter(map_lgl(parsed_crew, ~is.data.frame(.x) && nrow(.x) > 0)) |>
  unnest(parsed_crew, names_repair = "unique") |>
  rename(
    person_id = id, 
    original_crew_credit_id = credit_id, 
    job_name = job,
    person_name_crew = name
  ) |>
  select(movie_id, original_crew_credit_id, person_id, person_name_crew, job_name, department, gender) |>
  arrange(movie_id, original_crew_credit_id, person_id, job_name)

# Process the 'cast' JSON column
credits_cast_raw <- raw_credits |>
  select(movie_id, cast_json = cast) |>
  mutate(parsed_cast = map(cast_json, parse_json_safely)) |>
  filter(map_lgl(parsed_cast, ~is.data.frame(.x) && nrow(.x) > 0)) |>
  unnest(parsed_cast, names_repair = "unique") |>
  rename(
    person_id = id, 
    original_cast_credit_id = credit_id, 
    character_name_raw = character,
    person_name_cast = name
  ) |>
  select(movie_id, original_cast_credit_id, person_id, person_name_cast, character_name_raw, gender, order, cast_id) |>
  arrange(movie_id, original_cast_credit_id, person_id, order)

cat("--- Preliminary credits_crew_raw and credits_cast_raw processed ---\n")
cat("Head of credits_crew_raw:\n")
head(credits_crew_raw)
cat("\nHead of credits_cast_raw:\n")
head(credits_cast_raw)
```	

## Person
```{r}
person_from_crew <- credits_crew_raw |>
  distinct(person_id, person_name = person_name_crew)

person_from_cast <- credits_cast_raw |>
  distinct(person_id, person_name = person_name_cast)

person <- bind_rows(person_from_crew, person_from_cast) |>
  distinct(person_id, person_name) |>
  filter(!is.na(person_id), !is.na(person_name), person_name != "") |>
  # The ERD shows Person.id as INT. TMDB person_id is already an INT.
  # We will use person_id from TMDB as the primary key 'id' for the Person table.
  rename(id = person_id, name = person_name) |>
  arrange(id)

cat("\n--- Person created ---\n")
head(person)
```

## CrewRole
```{r}
roles_from_crew_jobs <- credits_crew_raw |>
  distinct(role_name = job_name)

# Add "Actor" role for cast members, as they are part of the crew concept here.
actor_role_tibble <- tibble(role_name = "Actor")

crew_role <- bind_rows(roles_from_crew_jobs, actor_role_tibble) |>
  distinct(role_name) |>
  filter(!is.na(role_name), role_name != "") |>
  arrange(role_name) |>
  mutate(id = row_number()) |> # ERD shows CrewRole.id as INT PK
  select(id, role = role_name) |> # ERD shows CrewRole.role VARCHAR
  arrange(id)

cat("\n--- CrewRole created ---\n")
head(crew_role)
```

## Character
```{r}
character <- credits_cast_raw |>
  distinct(character_name = character_name_raw) |>
  filter(!is.na(character_name), character_name != "") |>
  arrange(character_name) |>
  mutate(id = row_number()) |> # ERD shows Character.id as INT PK
  select(id, character = character_name) |> # ERD shows Character.character VARCHAR
  arrange(id)

cat("\n--- Character created ---\n")
head(character)
```
```

## Crew
DE AQUI PARA ABAJO NO ESTA TERMINADO. HASTA CHARACTER TODO OK, REVISADO POR RUBEN
```{r}
# Get crewRole_id for "Actor"
actor_role_id_value <- crew_role |> 
  filter(role == "Actor") |> 
  pull(id)

# If "Actor" role wasn't found (e.g., if crew_role was empty or "Actor" wasn't in jobs),
# this might be empty. Ensure it's handled.
if(length(actor_role_id_value) == 0){
  if (!("Actor" %in% crew_role$role)) {
    new_id_for_actor <- if (nrow(crew_role) > 0) max(crew_role$id) + 1 else 1
    crew_role <- bind_rows(crew_role, tibble(id = new_id_for_actor, role = "Actor"))
    actor_role_id_value <- new_id_for_actor
    cat("--- 'Actor' role dynamically added to crew_role with id:", actor_role_id_value, "---\n")
  } else {
     # This case should ideally not be reached if logic is sound
     actor_role_id_value <- crew_role |> filter(role == "Actor") |> pull(id)
  }
}


# Combine all credit entries (crew and cast) to generate a unique crew_id (INT)
# Storing original_credit_id to link with characters later if needed for cast.
combined_credits_for_crew_table <- bind_rows(
  # Crew members
  credits_crew_raw |>
    left_join(crew_role, by = c("job_name" = "role")) |>
    filter(!is.na(id)) |> # Ensure the role was found in crew_role
    select(
      original_credit_id_str = original_crew_credit_id,
      movie_id,
      person_id, # This is Person.id
      crewRole_id = id, # This is CrewRole.id
      type = "crew_member"
    ),
  # Cast members (as "Actor" crew)
  credits_cast_raw |>
    mutate(crewRole_id = actor_role_id_value) |>
    select(
      original_credit_id_str = original_cast_credit_id,
      movie_id,
      person_id, # This is Person.id
      crewRole_id, # This is CrewRole.id for "Actor"
      character_name = character_name_raw, # Keep for CrewCharacter
      type = "cast_member"
    )
) |>
  # The original_credit_id_str is unique for a (movie, person, role/character)
  # We need a new integer crew_id for the Crew table as per ERD
  distinct(original_credit_id_str, movie_id, person_id, crewRole_id, .keep_all = TRUE) |>
  arrange(movie_id, person_id, crewRole_id) |>
  mutate(crew_id = row_number()) # This is the new PK for Crew table (Crew.crew_id INT)

# Create the final Crew table
crew <- combined_credits_for_crew_table |>
  select(
    crew_id, # PK
    movie_id, # FK to Movie
    person_id, # FK to Person
    crewRole_id # FK to CrewRole
  ) |>
  arrange(crew_id)

cat("\n--- Crew created ---\n")
head(crew)
```

## CrewCharacter
```{r}
crew_character <- combined_credits_for_crew_table |>
  filter(type == "cast_member") |> # Only cast members play characters
  filter(!is.na(character_name), character_name != "") |>
  left_join(character_table, by = c("character_name" = "character")) |> # Join to get Character.id
  filter(!is.na(id)) |> # Ensure the character was found in character_table
  select(
    crew_id, # FK to Crew.crew_id
    character_id = id # FK to Character.id
  ) |>
  distinct() |>
  arrange(crew_id, character_id)

cat("\n--- CrewCharacter created ---\n")
head(crew_character)
```


La parte de `User`, `UserRole` etc no está en los csv. Por eso no hay extracción de datos de esos datasets.